An Aliasing-Tolerant Parallel Programming Model

### 2012-11-11 doc/types-pldi2013/alt_example.tex
Working on figures to make the prettier and easier to understand

The most important decision in any Legion program is how to partition
program data.  An ideal partitioning depends on many factors,
including the shape of data structures, the input, and the number of
partitions, which usually varies with the target machine.  To allow
the necessary flexibility, in Legion partitioning is simply programmed,
rather than expressed in a more structured and restricted
sublanguage.  In the example we invoke METIS\cite{Metis98}, a dynamic graph
partitioning library called (see Listing~\ref{lst:metis}).

\begin{lstlisting}[label={lst:metis},caption={Partition Computation}]
type NodeList<rl,rn>       = < CircuitNode@rn, node_list<rl,rn>@rl >
type WireList<rl,rw,rn,rg>= < CircuitWire<rn,rg>@rw, wire_list<rl,rw,rn,rg>@rl >
function extern_metis[rl,rn,rw](node_list : NodeList<rl,rn>@rl,
          wires_list : WireList<rl,rw,rn,rn>@rl), reads(rl,rn,rw), writes(rn) : bool;
\end{lstlisting}

@@

Note that there are two alternatives to using multiple partitions of the same data, both of which avoid introducing aliasing.
We could use a single partition with $2^n$ subregions, one for each
possible case of sharing; alternatively, all pieces could just use the
{\tt all\_nodes} region to access their ghost nodes.  Neither option
is attractive: the former significantly complicates programming, the
latter greatly increases data movement required at
runtime.

Programmers express partitions using {\em colorings},
maps from region elements to integers (``colors'').
Listing~\ref{lst:coloring} shows the function {\tt create\_piece\_coloring}
that constructs a coloring from the piece IDs of a list of nodes.
%The coloring of wires is similar, with each wire
%using the color of its first endpoint.  
For aliased partitions such as the ghost nodes, a {\em multicoloring}
is used, which assigns each region element (possibly) multiple colors.  
%(Multicolorings are
%implemented in Core Legion with separate colorings for each color.)
\begin{lstlisting}[label={lst:coloring},caption={Coloring Construction}]
function create_piece_coloring[rl,rn](node_list: NodeList<rl,rn>@rl),
            reads(rl,rn) : coloring(rn) = 
    if isnull(node_list) then
        newcolor(rn) 
    else
        let list_elem : NodeList<rl,rn> = read(node_list) in
        let part_coloring : coloring(rn) = create_node_coloring(list_elem.2) in
        let node : CircuitNode = read(list_elem.1) in
            color(part_coloring, list_elem.1, node.5)
\end{lstlisting}

@@

\begin{lstlisting}[label={lst:rr},caption={Region Relationship Example}]
type CircuitPiece<rl,rw,rn> = rr[rpw,rpn,rg]
                            < WireList<rl,rpw,rpn,rg>@rl, NodeList<rl,rpn>@rl >         
                            where rpn #$\le$# rn and rg #$\le$# rn and rpw #$\le$# rw and
                                  rn * rw and rl * rn and rl * rw
\end{lstlisting}

The Legion type system verifies properties of region relationships when packing.  When
a region relationship is unpacked later, fresh names are given to the regions that were
captured by the region relationship, and the properties that are known to hold for these
regions are re-introduced to the type environment.
In contrast, region access privileges cannot be packed in a region
relationship---privileges belong to functions. When a function unpacks
a region relationship it must already hold privileges for the unpacked regions it wishes to access.

>>>>

The most important decision in any Legion program is how to partition
program data.  An ideal partitioning depends on many factors,
including the shape of data structures, the input, and the number of
partitions, which usually varies with the target machine.  To allow
the necessary flexibility, in Legion partitioning is simply programmed,
rather than expressed in a more structured and restricted
sublanguage.  In the example we invoke METIS\cite{Metis98}, a 
third-party dynamic graph partitioning library (see Listing~\ref{lst:metis}).

\begin{lstlisting}[label={lst:metis},caption={Partition Computation}]
type NodeList<rl,rn>       = < CircuitNode@rn, NodeList<rl,rn>@rl >
type WireList<rl,rw,rn,rg>= < CircuitWire<rn,rg>@rw, WireList<rl,rw,rn,rg>@rl >
function extern_metis[rl,rn,rw](node_list : NodeList<rl,rn>@rl,
          wires_list : WireList<rl,rw,rn,rn>@rl), reads(rl,rn,rw), writes(rn) : bool
\end{lstlisting}

@@

Note there are two alternative approaches to using multiple partitions,
both of which avoid introducing aliasing:
first, we could create a single partition with $2^n$ subregions, one for each
possible case of sharing; second, tasks for each piece could use the
{\tt all\_nodes} region to access their ghost nodes.  Neither option
is attractive: the former significantly complicates programming, the
latter greatly increases runtime data movement and would limit parallelism.

Programmers express partitions using {\em colorings},
maps from region elements to integers (``colors'').
Listing~\ref{lst:coloring} shows the function {\tt create\_piece\_coloring}
that constructs a coloring from the piece IDs of a list of nodes.
%The coloring of wires is similar, with each wire
%using the color of its first endpoint.  
For aliased partitions such as the ghost nodes, a {\em multicoloring}
is used, which assigns each region element to one or more colors.  
%(Multicolorings are
%implemented in Core Legion with separate colorings for each color.)
\begin{lstlisting}[label={lst:coloring},caption={Coloring Construction}]
function create_piece_coloring[rl,rn](node_list: NodeList<rl,rn>@rl),
            reads(rl,rn) : coloring(rn) = 
    if isnull(node_list) then
        newcolor(rn) 
    else                                 -- tuple fields accessed by .(field number)
        let list_elem : NodeList<rl,rn> = read(node_list) in
        let part_coloring : coloring(rn) = create_node_coloring(list_elem.2) in
        let node : CircuitNode = read(list_elem.1) in
            color(part_coloring, list_elem.1, node.5)
\end{lstlisting}

@@

\begin{lstlisting}[label={lst:rr},caption={Region Relationship Example}]
type CircuitPiece<rl,rw,rn> = rr[rpw,rpn,rg]  -- rr is keyword for region relationship 
                            < WireList<rl,rpw,rpn,rg>@rl, NodeList<rl,rpn>@rl >         
                            where rpn #$\le$# rn and rg #$\le$# rn and rpw #$\le$# rw and
                                  rn * rw and rl * rn and rl * rw
\end{lstlisting}

The Legion type system verifies properties of region relationships when packing.  When
a region relationship is unpacked later, fresh names are given to the regions that were
captured by the region relationship (passed in {\tt []} parameters), and the properties that are known to hold for these
regions are re-introduced to the type environment.
In contrast, region access privileges cannot be packed in a region
relationship---privileges belong to functions. When a function unpacks
a region relationship it must already hold privileges for the unpacked regions it wishes to access.

### 2012-11-11 doc/types-pldi2013/intro.tex
Working on figures to make the prettier and easier to understand

In this work, we present the static and dynamic semantics of Legion \cite{Legion12},
a parallel programming model that supports multiple, dynamic data partitions and is
able to efficiently reason about potential aliasing.
Legion's {\em logical regions} are an abstraction capturing {\em locality} 
(data that will be used together, and therefore should be colocated) 
and {\em independence} (disjoint data that may be operated on in parallel).  
Logical regions have properties that allow programmers to communicate information
about the structure of program data to the Legion implementation:
\begin{itemize}
\item  logical regions are first-class values in Legion
and may be dynamically allocated and stored in data structures;

@@

To support additional parallelism in the presence of aliased data,
Legion provides {\em coherence} modes at the granularity of regions,
which allow relaxed constraints on task execution order, even
in cases where tasks access the same region.  The programmer may
specify that tasks conflicting on a region run in
program order (the default), {\em atomically} with respect to each
other, or even {\em simultaneously} if the application has alternative
methods for ensuring determinism or can tolerate non-deterministic
behavior in that region.

>>>>

In this work, we present the static and dynamic semantics for Legion \cite{Legion12},
a parallel programming model that supports multiple, dynamic data partitions and is
able to efficiently reason about potential aliasing.
Legion's {\em logical regions} are an abstraction that captures {\em locality} 
(data that will be used together, and therefore should be colocated) 
and {\em independence} (disjoint data that may be operated on in parallel).  
Logical regions have properties that allow programmers to communicate information
about the structure of program data to the Legion implementation:
\begin{itemize}
\item  logical regions are first-class values in Legion
and may be dynamically allocated and stored in data structures;

@@

To support additional parallelism when tasks access aliased regions,
Legion provides {\em coherence} modes at the granularity of regions, 
which allow relaxed constraints on task execution order.  The programmer may
specify that tasks accessing aliased regions run in
program order (the default), {\em atomically} with respect to each
other, or even {\em simultaneously} if the application has alternative
methods for managing non-deterministic
behavior in that region.
